---
title: "HuYuDataInsight Zhaowei Cai"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Loading packages
```{r}
library(tseries)
library(forecast)
library(stringr)
library(PerformanceAnalytics)
library(xts)
```

# Question 1
```{r}
set.seed(123)
#create a time series with 531 observations and first element is 0
ts_sim <- arima.sim(list(order = c(1,1,0), ar=0.65), n = 530)
```


```{r}
plot(ts_sim, col=4, las=1)
abline(v=c(1, 501, 531), lty="dotted", col=4)
```

# Question 2

Used 501 observations for myts because ts_sim[1] is always 0 and ts_sim has 531 observations
```{r}
myts = subset(ts_sim, subset=rep(c(TRUE, FALSE), times=c(501, 30)))
```

## Step 1: visualize myts
```{r}
plot.zoo(myts, col=4, las=1, main="Time Series")
```
## Step 2: unit root test (augmented Dickey-Fuller) of myts
```{r}
adf.test(myts, alternative = 'stationary')
```
P-value greater than 0.05, not reject H0, and it is not stationary.

## Step 3: differentiate myts, creating mydts
```{r}
mydts = diff(myts)
```

## Step 4: unit root test (augmented Dickey-Fuller) of mydts
```{r}
adf.test(mydts, alternative = 'stationary')
```
P-value less than 0.05, not reject H0, and it is stationary.

## Step 5: identify lags for mydts
```{r}
par(mfrow=c(1,2), mar=c(5,4,3,3))
acf(mydts)
pacf(mydts)
```
ACF decreases slowly, but PACF shows that it is an AR(1) (lag=1 is relevant only).

## Step 6: train the model with auto.arima for mydts
```{r}
fit_mydts = auto.arima(mydts, max.p=3, max.q=3, ic="aicc",
                       seasonal=FALSE, stationary=TRUE, lambda=NULL, 
                       stepwise=FALSE, approximation=FALSE
                       )
summary(fit_mydts)
```

## Step 7: fit the original time series, i.e. myts

```{r}
fit_myts = arima(myts, c(1, 1, 0))
summary(fit_myts)
```

Or can directly fit the original time series
```{r}
fit_myts2 = auto.arima(myts)
summary(fit_myts2)
```

# Question 3

## Question 3(a)
```{r}
forecast_myts = forecast(fit_myts, h=30, level=0.95)
plot(forecast_myts, col=4, las=1)
abline(v=c(1, 501, 531), lty="dotted", col=4)
lines(502:531, ts_sim[502:531], lty="dotted", col="red")
# red is observation and blue is prediction
```

## Question 3(b)
```{r}
# since it is one step ahead prediction, so we need use for loop
pred_df <- data.frame(NULL)
for(t in 502:531){
  pred_onestep <- forecast(ts_sim[1:t], h=1, level=0.95, model = fit_myts)
  pred_df <- rbind(pred_df, data.frame(mean = pred_onestep$mean[1], lower = pred_onestep$lower[1], upper = pred_onestep$upper[1]))
}
```

```{r}
plot(ts_sim, col=4, las=1)
abline(v=c(1, 501, 531), lty="dotted", col=4)
lines(502:531, pred_df$mean, col = 'red')
lines(502:531, pred_df$lower, col = 'green')
lines(502:531, pred_df$upper, col = 'green')
legend(40, 40, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```
make the plot more clear
```{r}
plot(491:531, ts_sim[491:531], col=4, las=1, type = 'l', ylim = c(25,50))
abline(v=c(501, 531), lty="dotted", col=4)
lines(502:531, pred_df$mean, col = 'red')
lines(502:531, pred_df$lower, col = 'green')
lines(502:531, pred_df$upper, col = 'green')
legend(490, 50, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```

# Question 4

Generate AR(1) model data
```{r}
set.seed(123)
ts_sim = arima.sim(list(ar=0.65),n=530)
```
add trend in the data
```{r}
ts_sim=ts_sim + 0.3 + 0.4*time(ts_sim)  
```
Generate plots
```{r}
plot(ts_sim, col=4, las=1)
abline(v=c(1, 500, 530), lty="dotted", col=4)
```

# Question 5

```{r}
myts = subset(ts_sim, subset=rep(c(TRUE, FALSE), times=c(500, 30)))
```
Step 1: visualize myts
```{r}
plot.zoo(myts, col=4, las=1, main="Time Series")
```
Fit trend part
```{r}
time = time(myts)
reg=lm(myts~time)
summary(reg)
```

```{r}
plot(as.vector(time(myts)), rstudent(reg), ylab="Standardized Res.", xlab="Time", type="o")
```

```{r}
z=rstandard(reg)
par(mfrow=c(1,2))
acf(z)
pacf(z)
```
Remove the trend part and fit the residuals
```{r}
newts=ts(residuals(reg))
```

Step 6: train the model with auto.arima for newts

```{r}
fit_newts = auto.arima(newts, max.p=3, max.q=3, ic="aicc",
                       seasonal=FALSE, stationary=TRUE, lambda=NULL, 
                       stepwise=FALSE, approximation=FALSE
                       )
summary(fit_newts)
```

# Question 6

## Question 6(a)
```{r}
prediction <- forecast(fit_newts, h=30, level=0.95)
pred_df <- data.frame(time = 501:530)
pred_df$mean <- prediction$mean + predict(reg, newdata = data.frame(time = 501:530))
pred_df$lower <- prediction$lower + predict(reg, newdata = data.frame(time = 501:530))
pred_df$upper <- prediction$upper + predict(reg, newdata = data.frame(time = 501:530))
```

```{r}
plot(ts_sim, col=4, las=1)
abline(v=c(1, 501, 530), lty="dotted", col=4)
lines(501:530, pred_df$mean, col = 'red')
lines(501:530, pred_df$lower, col = 'green')
lines(501:530, pred_df$upper, col = 'green')
legend(40, 150, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```
```{r}
plot(491:530, ts_sim[491:530], col=4, las=1, type = 'l')
abline(v=c(501, 531), lty="dotted", col=4)
lines(501:530, pred_df$mean, col = 'red')
lines(501:530, pred_df$lower, col = 'green')
lines(501:530, pred_df$upper, col = 'green')
legend(490, 210, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```
## Question 6(b)
Predict the residuls part
```{r}
# since it is one step ahead prediction, we will need to use the for loop

# generate the residuals
rests <- ts_sim - predict(reg, newdata = data.frame(time = time(ts_sim)))

pred_df <- data.frame(NULL)
for(t in 501:530){
  pred_onestep <- forecast(rests[1:t], h=1, level=0.95, model = fit_newts)
  pred_df <- rbind(pred_df, data.frame(mean = pred_onestep$mean[1], lower = pred_onestep$lower[1], upper = pred_onestep$upper[1]))
}
```

add predicted trend back
```{r}
pred_df$mean <- pred_df$mean + predict(reg, newdata = data.frame(time = 501:530))
pred_df$lower <- pred_df$lower + predict(reg, newdata = data.frame(time = 501:530))
pred_df$upper <- pred_df$upper + predict(reg, newdata = data.frame(time = 501:530))
```

plot pred, obs and CI

```{r}
plot(ts_sim, col=4, las=1)
abline(v=c(1, 501, 530), lty="dotted", col=4)
lines(501:530, pred_df$mean, col = 'red')
lines(501:530, pred_df$lower, col = 'green')
lines(501:530, pred_df$upper, col = 'green')
legend(40, 150, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```

```{r}
plot(491:530, ts_sim[491:530], col=4, las=1, type = 'l')
abline(v=c(501, 531), lty="dotted", col=4)
lines(501:530, pred_df$mean, col = 'red')
lines(501:530, pred_df$lower, col = 'green')
lines(501:530, pred_df$upper, col = 'green')
legend(490, 210, legend=c("Observations", "Prediction", "Bounds of CI"),col=c("blue", "red", "green"),lty = 1, cex=0.8)
```



